---
title: "Rosenberg factorial structure analysis"
output:
  html_notebook: default
  pdf_document: default
---

Individual groups analysis

1. seperation by countries
```{r}
ROSENBERG <- read.csv("C:/Users/march/Desktop/ROSENBERG.csv", header=FALSE, stringsAsFactors=FALSE)
colnames(ROSENBERG) <- c('gender','nation','ses1','ses2','ses3','ses4','ses5','ses6','ses7','ses8','ses9','ses10')
US <- ROSENBERG[ROSENBERG$nation ==1,]
CN <- ROSENBERG[ROSENBERG$nation ==2,]
JP <- ROSENBERG[ROSENBERG$nation ==3,]
```

2. import libraries
```{r}
library(lavaan)
```

3. build single model
```{r}
GSE <- 'gse =~ ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+ses10'

Bifactor <- 'pos =~ ses1+ses3+ses4+ses7+ses10
             neg =~ ses2+ses5+ses6+ses8+ses9'
### correlate uniqueness
cor.neg <- 'gse =~ ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+ses10
            ses2 ~~ ses5+ses6+ses8+ses9
            ses5 ~~ ses6+ses8+ses9
            ses6 ~~ ses8+ses9
            ses8 ~~ ses9'
cor.pos <- 'gse =~ ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+ses10
            ses1 ~~ ses3+ses4+ses7+ses10
            ses3 ~~ ses4+ses7+ses10
            ses4 ~~ ses7+ses10
            ses7 ~~ ses10'
cor.pos.neg <- 'gse =~ ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+ses10
                ses2 ~~ ses5+ses6+ses8+ses9
                ses5 ~~ ses6+ses8+ses9
                ses6 ~~ ses8+ses9
                ses8 ~~ ses9
                ses1 ~~ ses3+ses4+ses7+ses10
                ses3 ~~ ses4+ses7+ses10
                ses4 ~~ ses7+ses10
                ses7 ~~ ses10'
### method factor
neg.meth <- 'gse =~ ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+ses10
             neg =~ ses2+ses5+ses6+ses8+ses9
             gse ~~ 0*neg'
pos.meth <- 'gse =~ ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+ses10
             pos =~ ses1+ses3+ses4+ses7+ses10
             gse ~~ 0*pos'
pos.neg <- 'gse =~ ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+ses10
            neg =~ ses2+ses5+ses6+ses8+ses9
            pos =~ ses1+ses3+ses4+ses7+ses10
            gse ~~ 0*neg
            gse ~~ 0*pos
            neg ~~ 0*pos'
```


4. test model by nations
4.1.1 US general se
```{r}
US.fit <- cfa(GSE, data = US, estimator='ML')
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```

4.1.2 CN general se
```{r}
CN.fit <- cfa(GSE, data = CN, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```


4.1.3 JP general se
```{r}
JP.fit <- cfa(GSE, data = JP, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```

4.2.1 US Bifactor
```{r}
US.fit <- cfa(Bifactor, data = US, estimator='ML')
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```

4.2.2 CN Bifactor
```{r}
CN.fit <- cfa(Bifactor, data = CN, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```

4.2.3 JP Bifactor
```{r}
JP.fit <- cfa(Bifactor, data = JP, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```

4.3.1 US negative corr
```{r}
US.fit <- cfa(cor.neg, data = US, estimator='ML')
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```


4.3.2 CN negative corr
```{r}
CN.fit <- cfa(cor.neg, data = CN, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```


4.3.3 JP negative corr
```{r}
JP.fit <- cfa(cor.neg, data = JP, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```

### 
base on modification indices, the negative method effect model fit very well with JP & US population, for CN population, it indicates there some uniqueness among positive indicators need to be correlated

4.4.1 US positive corr
```{r}
US.fit <- cfa(cor.pos, data = US, estimator='ML')
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```


4.4.2 CN positive corr
```{r}
CN.fit <- cfa(cor.pos, data = CN, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```

4.4.3 JP positive corr
```{r}
JP.fit <- cfa(cor.pos, data = JP, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```

4.5.1 US positive & negative corr
```{r}
US.fit <- cfa(cor.pos.neg, data = US,std.lv=T, estimator='ML')
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```

 4.5.2 CN positive & negative corr
```{r}
CN.fit <- cfa(cor.pos.neg, data = CN,std.lv=T, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```
 
4.5.3 JP positive & negative corr
```{r}
JP.fit <- cfa(cor.pos.neg, data = JP, std.lv=T, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```

### 
pos & neg corr not converge

4.6.1 US negative method
```{r}
US.fit <- cfa(neg.meth, data = US, estimator='ML',std.lv=T)
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```


4.6.2 CN negative method
```{r}
CN.fit <- cfa(neg.meth, data = CN,std.lv=T, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```


4.6.3 JP negative method
```{r}
JP.fit <- cfa(neg.meth, data = JP, std.lv=T, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```

4.7.1 US positive method
```{r}
US.fit <- cfa(pos.meth, data = US, estimator='ML')
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```


4.7.2 CN positive method
```{r}
CN.fit <- cfa(pos.meth, data = CN, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```


4.7.3 JP positive method
```{r}
JP.fit <- cfa(pos.meth, data = JP, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```


4.8.1 US pos neg method
```{r}
US.fit <- cfa(pos.neg, data = US, estimator='ML')
summary(US.fit, fit.measures=T)
US.mi <- modindices(US.fit)
US.mi[US.mi$mi > 10,]
```

4.8.2 CN pos neg method
```{r}
CN.fit <- cfa(pos.neg, data = CN, estimator='ML')
summary(CN.fit, fit.measures=T)
CN.mi <- modindices(CN.fit)
CN.mi[CN.mi$mi > 10,]
```


4.8.3 JP pos neg method
```{r}
JP.fit <- cfa(pos.neg, data = JP, estimator='ML')
summary(JP.fit, fit.measures=T)
JP.mi <- modindices(JP.fit)
JP.mi[US.mi$mi > 10,]
```


MUltigroup Analysis: Measurement Equivalence

Configural Invariance

5.1 negative uniqueness correlated

```{r}
Mi.Fit <- cfa(cor.neg, data=ROSENBERG, group= 'nation', std.lv=T)
summary(Mi.Fit, fit.measures =T)
Mi.Fit.mi <- modificationindices(Mi.Fit)
Mi.Fit.mi[Mi.Fit.mi$mi >10,]
```


5.2 negative method effect
```{r}
Mi.Fit <- cfa(neg.meth, data=ROSENBERG, group= 'nation', std.lv=T)
summary(Mi.Fit, fit.measures =T)
Mi.Fit.mi <- modificationindices(Mi.Fit)
Mi.Fit.mi[Mi.Fit.mi$mi >10,]
```


5.3 negative & positive method
```{r}
Mi.Fit <- cfa(pos.neg, data=ROSENBERG, group= 'nation', std.lv=T)
summary(Mi.Fit, fit.measures =T)
Mi.Fit.mi <- modificationindices(Mi.Fit)
Mi.Fit.mi[Mi.Fit.mi$mi >10,]
```

Weak Invariance

6. set item 10 as refer indicator, based on results from factor ratio test

```{r}
cor.neg <- 'gse =~ NA*ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+1*ses10
            ses2 ~~ ses5+ses6+ses8+ses9
            ses5 ~~ ses6+ses8+ses9
            ses6 ~~ ses8+ses9
            ses8 ~~ ses9'
neg.meth <- 'gse =~ NA*ses1+ses2+ses3+ses4 +ses5+ses6+ses7+ses8+ses9+1*ses10
             neg =~ ses2+ses5+ses6+ses8+ses9
             gse ~~ 0*neg'
pos.neg <- 'gse =~ NA*ses1+ses2+ses3+ses4 +ses5+ses6+1*ses7+ses8+ses9+ses10
            neg =~ ses2+ses5+ses6+ses8+ses9
            pos =~ ses1+ses3+ses4+ses7+ses10
            gse ~~ 0*neg
            gse ~~ 0*pos
            neg ~~ 0*pos'
```


6.1 Correlated negative indicators
```{r}
Wi.fit <- cfa(cor.neg, data= ROSENBERG, group='nation',group.equal='loadings')
summary(Wi.fit, fit.measures =T)
```


6.2 method effect
```{r}
Wi.fit <- cfa(neg.meth, data= ROSENBERG, group='nation',group.equal='loadings',
              group.partial = c('neg =~ses2',
                                'neg =~ses5',
                                'neg =~ses6',
                                'neg =~ses8',
                                'neg =~ses9'))
summary(Wi.fit, fit.measures =T)
```

6.3 Positive & Negative method effect
```{r}
Wi.fit <- cfa(pos.neg, data= ROSENBERG, group='nation',group.equal='loadings',
              group.partial = c('neg =~ses2',
                                'neg =~ses5',
                                'neg =~ses6',
                                'neg =~ses8',
                                'neg =~ses9',
                                'pos =~ses1',
                                'pos =~ses3',
                                'pos =~ses4',
                                'pos =~ses7',
                                'pos =~ses10'))
summary(Wi.fit, fit.measures =T)
```

No converge

Strict Invariance

7.1 correlated negative indicator
```{r}
Si.fit <-cfa(cor.neg, data= ROSENBERG, group='nation',group.equal=c('loadings','intercepts'))
summary(Si.fit, fit.measures =T)
```


7.2 Negative method
```{r}
Si.fit <-cfa(neg.meth, data= ROSENBERG, group='nation',group.equal=c('loadings','intercepts'))
summary(Si.fit, fit.measures =T)
```


7.3 Negative and Positive method
```{r}
Si.fit <- cfa(pos.neg, data= ROSENBERG, group='nation',group.equal=c('loadings','intercepts'),
              group.partial = c('neg =~ses2',
                                'neg =~ses5',
                                'neg =~ses6',
                                'neg =~ses8',
                                'neg =~ses9',
                                'pos =~ses1',
                                'pos =~ses3',
                                'pos =~ses4',
                                'pos =~ses7',
                                'pos =~ses10'))
summary(Si.fit, fit.measures =T)
```


























